---
title: <span style="color:#e8b22a">Tidy Time Series Graphics with R</span>
subtitle: "IASSL Workshop"
author: "Dr. Priyanga D. Talagala, University of Moratuwa"
#institute: "RStudio, PBC"
date: <span style="color:#3be3dd"> 21-25, February, 2022</span>
output:
  xaringan::moon_reader:
    #css: [default, kunoichi, chocolate-fonts]
    css: [ "default", "rladies", "rladies-fonts", "xaringan-themer.css"]
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: "%current%" 
---
```{r setup, include=FALSE}
library(knitr)
options(htmltools.dir.version = FALSE)
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  cache.path = 'cache/',
  fig.align = 'center',
  cache = TRUE
)

library(tidyverse)
library(patchwork)

```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent(
   base_color = "#025859",
   link_color = "#1148ed",
   title_slide_background_color = "#025859",
  inverse_background_color = "#17ffa2",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono"),
  title_slide_text_color = "#3be3dd",
  inverse_header_color = "#500252",
   inverse_text_color = "#041a5c",
   header_background_text_color = "#ffea00",
  base_font_size = "20px",
  text_font_size = "1.5rem",
  )
```

### Main packages required

```{r echo=TRUE}
# Data manipulation and plotting functions
library(tidyverse)

# Data for "Forecasting: Principles and Practice" (3rd Edition)
library(fpp3)
```


- `tsibble` Package:  Time series manipulation 
- `tsibbledata` Package: Tidy time series data 
- `feasts` Package: Time series graphics and statistics 
- `fable` Package: Forecasting functions


tidyverts: Tidy tools for time series: https://tidyverts.org/
---
## `tsibble` objects

-  A tsibble allows storage and manipulation of multiple time series in R.
-  It contains:

   – An index: time information about the observation
   
   – Measured variable(s): numbers of interest
   
   – Key variable(s): optional unique identifiers for each series 
   
   – It works with tidyverse functions.
---

## The `tsibble` index

.pull-left[
```{r echo=TRUE}
set.seed(1)
ts <- tsibble(t = seq(36),
              y = rnorm(36),
              index = t) #<<
ts
```



]
--
.pull-right[

```{r echo=TRUE}
mydata <- tsibble(
  year = 2016:2020,
  y = c(123, 39, 78, 52, 110),
  index = year) #<<

mydata

```
]

---
## `tibble` vs `tsibble`

.pull-left[

```{r echo=TRUE}
#tibble
mytibble <- tibble(
  date = as.Date("2017-01-01") + 0:10,
  y = c(123, 39, 78, 52, 110, 59,
        78, 67, 67, 80, 90)) 
mytibble
```

]
--
.pull-right[

```{r echo=TRUE}
# Converting to a tsibble
mytsibble <- mytibble %>% 
  as_tsibble(index = date) #<<
mytsibble
```
]
---

.pull-left[

```{r echo=TRUE}
mytsibble
```

]
--
.pull-right[

```{r echo=TRUE}
mytsibble %>% autoplot(y)
```
]

---

.pull-left[
### Working with tsibble objects

Example: Quarterly Australian Electricity Production

```{r echo=TRUE}
elec <- aus_production %>%
  select(Quarter, Electricity) %>% 
  filter(year(Quarter) >= 1992)
elec
```

]
--
.pull-right[
```{r echo=TRUE, fig.height=5}
elec %>%
  autoplot(Electricity) #<<
```
]
---

.pull-left[
## Seasonal plots

```{r echo=TRUE}
elec <- aus_production %>%
  select(Quarter, Electricity) %>% 
  filter(year(Quarter) >= 1992)
elec
```

]
--
.pull-right[
```{r echo=TRUE, fig.height=6}
elec %>% 
  gg_season(Electricity,  #<<
            labels="both") #<<
```
]
---


.pull-left[
## Seasonal plots

```{r echo=TRUE, fig.height=6}
elec %>% 
  gg_season(Electricity, 
            labels="left") #<<
```

]
.pull-right[
- Data plotted against the individual “seasons” in which the data were observed. (In this case a “season” is a month.)

- Something like a time plot except that the data from each season are overlapped.

- Enables the underlying seasonal pattern to be seen more clearly, and also allows any substantial departures from the seasonal pattern to be easily identified.
]
---
```{r echo=TRUE, fig.height=6}
elec %>% 
  gg_season(Electricity, labels="left") + 
  ylab("Electricity production in gigawatt hours") + #<<
  ggtitle("Quarterly Australian Electricity Production") #<<
```

---

.pull-left[
## Seasonal subseries plots

- Data for each season collected together in time plot as separate time series.

- Enables the underlying seasonal pattern to be seen clearly, and changes in seasonality over time to be
visualized.

]

--
.pull-right[

```{r echo=TRUE}
elec %>% gg_subseries(Electricity)
```

]
---
## Multiple seasonal periods

Half-hourly electricity demand for Victoria, Australia

```{r echo = TRUE}
vic_elec
```


---
```{r echo=TRUE, fig.width= 11}
vic_elec %>% 
  autoplot(Demand)
```
---

```{r echo=TRUE,  fig.width= 11}
vic_elec %>% 
  gg_season(Demand) #<<
```
---
```{r echo=TRUE,  fig.width= 11}
vic_elec %>% 
  gg_season(Demand, period = "month") #<<
```
---
```{r echo=TRUE,  fig.width= 11}
vic_elec %>%
  gg_season(Demand, period = "week") #<<
```
---

```{r echo=TRUE,  fig.width= 11}
vic_elec %>%
  gg_season(Demand, period = "day") #<<
```

---

```{r echo=TRUE,  fig.width= 11}
vic_elec %>%
  gg_season(Demand, period = "day") +
  theme(text = element_text(size=14)) #<<
```
---
.pull-left[
```{r echo=TRUE}
# install.packages("devtools")
# devtools::install_github("thiyangt/covid19srilanka")
library(covid19srilanka)
district.wise.cases
```

]

--
.pull-right[

```{r echo=TRUE}
covidsl <- district.wise.cases %>%
  as_tsibble(index = Date, #<<
             key = District) #<<

covidsl
```

]


---


```{r echo=TRUE, fig.width=15}
p <- covidsl %>% autoplot(Count)
p
```

---
```{r echo=TRUE, fig.width=15 }
plotly::ggplotly(p)
```

---

```{r echo=TRUE}
library(coronavirus)
head(coronavirus)
```

---

.pull-left[
```{r echo=TRUE}
covidsl <- coronavirus %>%
  filter(country == "Sri Lanka") %>%
  select(date, type, cases)

head(covidsl)
```

]

--
.pull-right[

```{r echo=TRUE}
covidsl <- covidsl %>%
  as_tsibble(index = date, #<<
             key = type ) #<<

head(covidsl)
```

]

---

```{r echo=TRUE, fig.width=11}
p <- covidsl %>% autoplot(cases)
p
```

---

```{r echo=TRUE, fig.width=11}
library(gganimate)
p +
  transition_reveal(along = date)
```

---
class: inverse, middle, center

```{r}
icon::fa("github")
icon::fa("twitter")

```
pridiltal and thiyangt

# Acknowledgements:

This work was supported in part by RETINA research lab funded by the OWSD, a program unit of United Nations Educational, Scientific and Cultural Organization (UNESCO).

## Key References 

- Forecasting: Principles and Practice (3rd ed), Rob J Hyndman and George Athanasopoulos https://otexts.com/fpp3/


All rights reserved by Thiyanga S. Talagala and Priyanga D Talagala



